Hybrid recurrent model for regime switching nonstationary time series

### 2021-10-03 iclr2022/hybridmodel_iclr.tex
update results and few corrections

We provide the first dataset dedicated to fashion trends forecasting. Our fashion dataset is built with a high throughput computer vision system analysing every day images shared on social networks. Aggregating detected clothes by trends, we create for the first time thousand of time series representing fashion trends life on social media. Our sequences have all 261 time steps, from 2015-01-05 to 2019-12-31 with weekly values and no missing walues. Each time series is normalized in order to remove social networks biases and their values lie between 0 and 1. The proposed fashion dataset contains $\numberts$ anonymized fashion trends for men and women, on 9 differents categories and 5 geozones. An overview of it can be found in Table~\ref{sample-table}.

With in-house fashion experts, we selected especially a collection of $\numberts$ fashion trends in order to represent finely the issues that face the fashion industry. A large part of our time series are quite stable, their seasons and evolutions are regular. However, some sequences show interesting behaviours with sudden changes. We call them emerging or declining trends depending of the change direction. A central point of our work is to accurately detect and forecast these trends.

@@

In addition to this large dataset we design for each sequence several weak signals.  Firstly, we create a specific fashion-oriented panel of micro-influencers. In theoretical fashion dynamics \citep{rogersdiffusion}, different categories of adopters follow a trend in succession, resulting in several adoption waves. With this specific panel, we aim at detecting the first waves announcing emerging trends or the collapse of other ones. By analyzing images shared by this specific panel on social medias, we create a first weak signal named \textit{fashion-forwards} for each of the previous trends. We also analyzed only a specific part of social media users depending of the number of followers. With three thresholds we created the following external signals : \textit{followers-low}, \textit{followers-mid} and \textit{followers-high}. Again, the motivation here is to detect behaviour gaps between these three segments that could announce bigger changes in the main signal. These extra signals are frequently sparse for micro trends and they often lack of interest for common fashion trends. However on several examples, they are fundamental to detect the future fashion evolution.

@@

As an example, an impressive emerging fashion trend is represented in Figure~\ref{fig:oneemergingtrend}, see other examples in Appendix~\ref{}. It shows the evolution of one shoes trends on social media and its linked \textit{fashion-forward} weak signal. During the first 3 years, common users and the panel of mode-influencers share the same behaviour. The trend start to boom in the mainstream people during the first month of 2018. However, we can detect early signals in the fashion-forward panel at the end of 2017. Early adopters started to change there behaviour and show more frequently the trend on social media. With fashion influence dynamics, the new tendency spreads to every part of social media users. Finally, after a pick reached at the end of 2018 for the fashion-influencers, a regular decrease appear. %They leave gradually the old trend to focus more on the latest or propose new ones.

@@

As benchmarks, we chose several well-known statistical methods and deep learning approaches. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels, tbats} , we compute for each time series a prediction with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. The \textit{snaive} forecasts are only the repetition of the last same past period. The \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. The \textit{stlm} approach uses a multiplicative decomposition and models the seasonality adjusted time series with an exponenantial smoothing model. The \textit{Thetam} model decomposes the original signal in $\theta$-lines, forecasts each one separately and recomposes them to produce the final forecast and  \textit{tbats} is a powerful statistical model using a trigonometrical seasonality modelization. Finally,  \textit{auto.arima} is the R implementation of ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citep{packageforecast}. As a deep learning approach, we considered a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52.

Against these models, we propose our two versions of our regime switching hybrid reccurent model (rshm) called respectively rshm-ets and rshm-tbats according to the per-time-series model choice.
%Against these models, we propose two versions of our regime switching hybrid recurrent model \textit{rshm}. The first one, that we name \textit{rshm-ets}, is the hybrid model with an additive exponential smoothing model as parametric model. For the second one, we use the TBATS approach as paramteric model and we named it \textit{rshm-tbats}. 
%To reach a higher accuracy level we also propose an ensembling (\textit{ensembling} of our hybrid approach using the FFORMA method created during the M4 competition, see \citep{fforma}). 
For our two versions, in a first time, we propose a version without our weak signal called \textit{rshm-ets} and \textit{rshm-tbats}. Then we evaluate our hybrid framework with the inclusion of our weak signals (ws) referred to as \textit{rshm-ets-ws} and \textit{rshm-tbats-ws}. In order to provide a fair comparison, we also train a \textit{lstm} with the weak signals named \textit{lstm-ws}.

@@

Detecting emerging and declining trends is a crucial issue for the fashion industry. A correct or incorrect prediction could lead to good returns or massive waste due to overstock or unsold clothes. In a given year, we define an increasing trend as a trend that does more than 10\% of growth in term of Year and Year. In the same way, we define a decreasing trend as a trend that declines of 10\% or more. We classify other trends as flat trends. On the test set, the year 2019, we have 1871 increasing trends, 2016 decreasing trends and 6113 flat trends. In addition to the accuracy metrics, we compare also the different methods on this classification task.

@@

>>>>

We provide the first dataset dedicated to fashion trends forecasting. Our fashion dataset is built with a high throughput computer vision system analysing every day images shared on social networks. Aggregating detected clothes by trends, we create for the first time thousands of time series representing fashion trends life on social media. Our sequences have all 261 time steps, from 2015-01-05 to 2019-12-31 with weekly values and no missing values. Each time series is normalized in order to remove social networks biases and their values lie between 0 and 1. The proposed fashion dataset contains $\numberts$ anonymized fashion trends for men and women, on 9 different categories and 5 geozones. An overview of it can be found in table 1.

With in-house fashion experts, we selected especially a collection of $\numberts$ fashion trends in order to represent finely the issues that face the fashion industry. A large part of our time series are quite stable, their seasons and evolutions are regular. However, some sequences show interesting behaviours with sudden changes. We call them emerging or declining trends depending on the change in direction. A central point of our work is to accurately detect and forecast these trends.

@@

In addition to this large dataset we design for each sequence several weak signals.  Firstly, we create a specific fashion-oriented panel of micro-influencers. In theoretical fashion dynamics \citep{rogersdiffusion}, different categories of adopters follow a trend in succession, resulting in several adoption waves. With this specific panel, we aim at detecting the first waves announcing emerging trends or the collapse of other ones. By analyzing images shared by this specific panel on social media, we create a first weak signal named \textit{fashion-forwards} for each of the previous trends. We also analyzed only a specific part of social media users depending on the number of followers. With three thresholds we created the following external signals : \textit{followers-low}, \textit{followers-mid} and \textit{followers-high}. Again, the motivation here is to detect behaviour gaps between these three segments that could announce bigger changes in the main signal.These extra signals are frequently sparse for micro trends and they often lack interest in common fashion trends. However, in several examples, they are fundamental to detect the future fashion evolution.

@@

As an example, an impressive emerging fashion trend is represented in Figure~\ref{fig:oneemergingtrend}, see other examples in Appendix~\ref{}. It shows the evolution of one shoe trend on social media and its linked \textit{fashion-forward} weak signal. During the first 3 years, common users and the panel of mode-influencers share the same behaviour. The trend started to boom in the mainstream people during the first month of 2018. However, we can detect early signals in the fashion-forward panel at the end of 2017. Early adopters started to change their behaviour and show more frequently the trend on social media. With fashion influence dynamics, the new tendency spreads to every part of social media users. Finally, after a peak reached at the end of 2018 for the fashion-influencers, a regular decrease appeared. %They leave gradually the old trend to focus more on the latest or propose new ones.

@@

As benchmarks, we chose several well-known statistical methods and deep learning approaches. Using the R package \texttt{forecast} and the Python packages \texttt{statsmodels, tbats} , we compute for each time series a prediction with the following methods: \textit{snaive}, \textit{ets}, \textit{stlm}, \textit{thetam}, \textit{tbats} and \textit{auto.arima}. \textit{snaive} forecasts are only the repetition of the last same past period. \textit{ets} model is an additive exponential smoothing with a level component and a seasonal component. \textit{stlm} approach uses a multiplicative decomposition and models the seasonally adjusted time series with an exponential smoothing model. \textit{Thetam} model decomposes the original signal in $\theta$-lines, predicts each one separately and recomposes them to produce the final forecast. \textit{tbats} is a powerful statistical model using a trigonometrical seasonality modelization. The last one \textit{auto.arima} is the R implementation of the ARIMA model with an automatic selection of the best parameters. A complete description and references for these models can be found in \citep{hyndman2020package}. As a deep learning approach, we considered a full LSTM (\textit{lstm}) neural network composed of 3 LSTM layers of shape 50 and a final Dense layer of shape 52.

Against these models, we propose two versions of our regime switching hybrid recurrent model (rshm) called respectively rshm-ets and rshm-tbats according to the per-time-series model choice.
%Against these models, we propose two versions of our regime switching hybrid recurrent model \textit{rshm}. The first one, that we name \textit{rshm-ets}, is the hybrid model with an additive exponential smoothing model as parametric model. For the second one, we use the TBATS approach as paramteric model and we named it \textit{rshm-tbats}. 
%To reach a higher accuracy level we also propose an ensembling (\textit{ensembling} of our hybrid approach using the FFORMA method created during the M4 competition, see \citep{fforma}). 
For our two versions, we propose a version without our weak signal called \textit{rshm-ets} and \textit{rshm-tbats}. Then we evaluate our hybrid framework with the inclusion of our weak signals (ws) referred to as \textit{rshm-ets-ws} and \textit{rshm-tbats-ws}. In order to provide a fair comparison, we also train a \textit{lstm} with the weak signals named \textit{lstm-ws}.

@@

Detecting emerging and declining trends is a crucial issue for the fashion industry. A correct or incorrect prediction could lead to good returns or massive waste due to overstock or unsold clothes. In a given year, we define an increasing trend as a trend that does more than 10\% of growth in terms of Year and Year.In the same way, we define a decreasing trend as a trend that declines by 10\% or more. We classify other trends as flat trends. On the test set, the year 2019, we have 1871 increasing trends, 2016 decreasing trends and 6113 flat trends. In addition to the accuracy metrics, we also compare the different methods on this classification task.

@@

\section{Additional results, explorations and examples.}
